---
title: "COVID-19 monitor"
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows
    logo: logo.png
    css: styles.css
runtime: shiny
---

<strong>Confirmed UK cases of COVID-19 as of `r renderText({format(daily_indicators$DateVal, '%d %B %Y')})`</strong><br/>Please note that the number of confirmed cases is likely to be much lower than the total number of cases because some people who report symptoms are not being tested. Further information about Coronavirus (COVID-19) can be found at: <a href="https://www.trafford.gov.uk/covid-19" target="_blank">trafford.gov.uk/covid-19</a>

```{r setup, include=FALSE}
library(flexdashboard) ; library(shiny) ; library(tidyverse) ; library(httr) ; library(readxl) ;  library(sf) ;  library(htmlwidgets) ; library(htmltools) ; library(leaflet) ; library(leaflet.extras) ; library(DT) ; library(scales)

# Data published by Public Health England 
# https://www.gov.uk/government/publications/covid-19-track-coronavirus-cases

# daily indicators
url <- "https://www.arcgis.com/sharing/rest/content/items/bc8ee90225644ef7a6f4dd1b13ea1d67/data"
GET(url, write_disk(tmp <- tempfile(fileext = ".xlsx")))
daily_indicators <- read_xlsx(tmp) %>% 
  mutate(DateVal = as.Date(DateVal, format = "%d/%m/%y"))

# cases by country
cases_by_country <- tibble(
  area_code = c("N92000002", "S92000003", "W92000004"),
  TotalCases = c(daily_indicators$NICases, daily_indicators$ScotlandCases, daily_indicators$WalesCases)
)

# number of daily confirmed cases
url <- "https://www.arcgis.com/sharing/rest/content/items/e5fd11150d274bebaaf8fe2a7a2bda11/data"
GET(url, write_disk(tmp <- tempfile(fileext = ".xlsx")))
daily_cases <- read_xlsx(tmp) %>% 
  mutate(DateVal = as.Date(DateVal, format = "%d/%m/%y"))

# UTLA cases table
utla_cases <- read_csv("https://www.arcgis.com/sharing/rest/content/items/b684319181f94875a6879bbc833ca3a6/data") %>% 
  select(area_code = GSS_CD, TotalCases) %>% 
  bind_rows(cases_by_country)

cases_by_area <- st_read("data/areas.geojson") %>% 
  left_join(utla_cases, by = "area_code") %>% 
  mutate(popup = str_c("<strong>", area_name, "</strong><br/>",
                       TotalCases, " cases") %>%
           map(HTML))
```

Row
-----------------------------------------------------------------------

### Total confirmed UK cases
```{r}
valueBox(comma(max(daily_cases$CumCases)))
```

### New confirmed UK cases on `r format(daily_indicators$DateVal, '%d %B %Y')`
```{r}
valueBox(comma(pull(filter(daily_cases, DateVal == max(DateVal)), CMODateCount)))
```

### Total UK deaths
```{r}
valueBox(comma(daily_indicators$TotalUKDeaths))
```

Row
-------------------------------------

### Cumulative confirmed cases
```{r}
renderPlot({
  ggplot(daily_cases, aes(x = DateVal, y = CumCases)) +
  geom_line(colour = "#57AACB", size = 1) +
  geom_point(colour = "#57AACB", size = 2) +
  scale_y_continuous(labels = comma, expand = c(0.005, 0.005), sec.axis = sec_axis(~ ., breaks = max(daily_cases$CumCases), labels = comma)) +
  scale_x_date(expand = c(0.005, 0.1), date_labels = "%d-%b") +
  geom_hline(yintercept = 0, size = 0.5, colour = "#212121") +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL,
       caption = "Source: Public Health England") +
  theme_minimal(base_size = 16) +
  theme(plot.margin = unit(rep(0.5, 4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 10, color = "grey50", hjust = 1, margin = margin(t = 15))) +
    expand_limits(y = c(0,max(daily_cases$CumCases)*1.01))
}, execOnResize = TRUE)
```

### Daily confirmed cases
```{r}
renderPlot({
  ggplot(daily_cases, aes(x = DateVal, y = CMODateCount)) +
  geom_col(fill = "#57AACB") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  scale_x_date(expand = c(0.005, 0.005), date_labels = "%d-%b") +
  scale_y_continuous(expand = c(0.005, 0.005), position = "right") +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL,
       caption = "Source: Public Health England") +
  theme_minimal(base_size = 16) +
  theme(plot.margin = unit(rep(0.5, 4), "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 10, color = "grey50", hjust = 1, margin = margin(t = 15))) 
}, execOnResize = TRUE)
```

Row
-------------------------------------

### Confirmed cases by country / UTLA
```{r}
leaflet() %>%
  setView(-3, 54.3, zoom = 5) %>% 
  addTiles(urlTemplate = "",
           attribution = '<a href="https://www.ons.gov.uk/methodology/geography/licences">Contains OS data © Crown copyright and database right (2020)</a> | Data: <a href="https://www.gov.uk/government/publications/covid-19-track-coronavirus-cases" target="_blank">Public Health England</a>') %>%
  addPolygons(data = cases_by_area, fillColor = "#EAEAEA", fillOpacity = 0.3, weight = 1, color = "#bdbdbd") %>% 
  addCircleMarkers(data = cases_by_area, lng = ~long, lat = ~lat, radius = ~sqrt(TotalCases), 
                   fillColor = "#57AACB", fillOpacity = 0.8, weight = 1, color = "#FFFFFF", opacity = 1,
                   label = ~popup) %>% 
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  onRender(
    "function(el, t) {var myMap = this;myMap._container.style['background'] = '#ffffff';}",
    paste0("function(el, x) {$('head').append(","\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'",");}"))
```

###
```{r}
renderDT({
  cases_by_area %>% 
    st_set_geometry(NULL) %>% 
    select(area_name, TotalCases) %>% 
    arrange(area_name) %>% 
    datatable(caption = "Confirmed cases by country / UTLA",
              extensions = 'Scroller',
              style = 'bootstrap', 
              class = 'table-bordered',
              fillContainer = TRUE,
              width = "100%",
              rownames = FALSE,
              escape = FALSE,
              options = list(
                dom = 't',
                #ordering = F,
                scrollX = TRUE,
                scrollY = 200,
                scroller = TRUE,
                columnDefs = list(list(className = 'dt-left', targets = "_all"))),
              colnames = c("Country / UTLA", "Cases"))
  })
```
