---
title: "COVID-19 UK monitor"
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows
    vertical_layout: fill
    logo: logo.png
    source_code: https://github.com/traffordDataLab/covid-19
    social: ["twitter"]
    css: styles.css
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard) ; library(shiny) ; library(tidyverse) ; library(httr) ; library(htmlwidgets) ; library(htmltools) ; library(zoo) ; library(scales) ; library(plotly) ; library(ggrepel) ; library(classInt) ; library(leaflet) ; library(leaflet.extras)

source("load_data.R")
```

Cases {data-navmenu="UK"}
=======================================================================

Row
-------------------------------------

### <strong>New confirmed UK cases</strong><br/><small>as of `r format(max(uk_data$Date), '%A %d %B %Y')`</small>
```{r}
valueBox(comma(pull(filter(uk_data, Date == max(Date)), NewCases)), color = "#39809E")
```

### <strong>Total confirmed UK cases</strong><br/><small>as of `r format(max(uk_data$Date), '%A %d %B %Y')`</small>
```{r}
valueBox(comma(max(uk_data$CumCases)), color = "#39809E")
```

Row
-------------------------------------

### Daily confirmed cases
```{r}
renderPlotly({
  
  daily_cases <- uk_data %>% 
    mutate(ma_cases = rollmean(NewCases, 7, align = "right", fill = NA))
  
  p <- ggplot() +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_col(data = daily_cases, aes(x = Date, y = NewCases, text = paste0("<b>", comma(NewCases, accuracy = 1), "</b> cases<br>", format(Date, "%d %B"))), fill = "#39809E", alpha = 0.6) +
    geom_line(data = daily_cases, aes(x = Date, y = ma_cases, text = "7 day rolling average"), colour = "#39809E", size = 1) +
    scale_x_date(expand = c(0.005, 0.005), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), labels = comma) +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank()) 
  
  ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>% 
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE))  
})
```

> **Notes**: 1) Confirmed cases are recorded up to 9am on `r format(max(uk_data$Date), '%A %d %B %Y')`. 2) There is likely to be a lag of a few days between initial testing and the reporting of test results. 3) The number of *confirmed* cases is likely to be much lower than the *total* number of cases because not all people who report symptoms of coronavirus are being tested.<br/>**Source**: [European Centre for Disease Prevention and Control](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide) from [data](https://www.gov.uk/guidance/coronavirus-covid-19-information-for-the-public) published by the UK [Department of Health and Social Care](https://www.gov.uk/government/organisations/department-of-health-and-social-care) and [Public Health England](https://www.gov.uk/government/organisations/public-health-england).
 
### Cumulative confirmed cases
```{r}
fillCol(flex = c(NA, 1),
        radioButtons("cases_scale", label = NULL, choices = c("Linear", "Logarithmic"), selected = "Linear", inline = TRUE),
        plotlyOutput("cases_plot", height = "100%")
        )

output$cases_plot <- renderPlotly({
  p <- ggplot(uk_data, aes(x = Date, y = CumCases, group = 1, text = paste0("<b>", comma(CumCases, accuracy = 1), "</b> cases<br>", format(Date, "%d %B")))) +
    geom_line(colour = "#39809E", size = 1) +
    scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank())
  
  if (input$cases_scale == "Linear") {
    
    p <- p + geom_hline(yintercept = 0, size = 0.3, colour = "#333333") + scale_y_continuous(expand = c(0.005, 0.005), labels = comma)
    
    } else {
      
      p <- p + geom_hline(yintercept = 1, size = 0.3, colour = "#333333") + scale_y_continuous(trans = "log10", expand = c(0.005, 0.005), labels = comma_format(accuracy = 1))
      
    }
  
  ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>% 
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE)) 
  
})
```

> **Notes**: Logarithmic scales represent an exponential curve as a straight line. The steeper the line the faster the rate of growth.

Deaths {data-navmenu="UK"}
=======================================================================

Row
-------------------------------------

### <strong>New confirmed UK deaths</strong><br/><small>as of `r format(max(uk_data$Date), '%A %d %B %Y')`</small>
```{r}
valueBox(comma(pull(filter(uk_data, Date == max(Date)), NewDeaths)), color = "#8D2313")
```

### <strong>Total confirmed UK deaths</strong><br/><small>as of `r format(max(uk_data$Date), '%A %d %B %Y')`</small>
```{r}
valueBox(comma(max(uk_data$CumDeaths)), color = "#8D2313")
```

Row
-------------------------------------

### Daily confirmed deaths
```{r}
renderPlotly({
  
  daily_deaths <- uk_data %>% 
    filter(Date >= "2020-03-06") %>% 
    mutate(ma_deaths = rollmean(NewDeaths, 7, align = "right", fill = NA))
  
  p <- ggplot() +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_col(data = daily_deaths, aes(x = Date, y = NewDeaths, text = paste0("<b>", comma(NewDeaths, accuracy = 1), "</b> deaths<br>", format(Date, "%d %B"))), fill = "#8D2313", alpha = 0.6) +
    geom_line(data = daily_deaths, aes(x = Date, y = ma_deaths, text = "7 day rolling average"), colour = "#8D2313", size = 1) +
    scale_x_date(expand = c(0.005, 0.005), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), labels = comma) +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank()) 
  
  ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>% 
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE)) 
})
```

> **Notes**: 1) Deaths are recorded up to 5pm on `r format(max(uk_data$Date)-1, '%A %d %B %Y')`. 2) The date of a death refers to the day the death was reported not the actual date of the death. 3) The number of deaths only includes those who tested positive for coronavirus. 4) The deaths reported cover all settings not just hospitals. <br/>**Source**: [European Centre for Disease Prevention and Control](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide) from [data](https://www.gov.uk/guidance/coronavirus-covid-19-information-for-the-public) published by the UK [Department of Health and Social Care](https://www.gov.uk/government/organisations/department-of-health-and-social-care) and [Public Health England](https://www.gov.uk/government/organisations/public-health-england).

### Cumulative confirmed deaths
```{r}
fillCol(flex = c(NA, 1),
        radioButtons("deaths_scale", label = NULL, choices = c("Linear", "Logarithmic"), selected = "Linear", inline = TRUE),
        plotlyOutput("deaths_plot", height = "100%")
        )

output$deaths_plot <- renderPlotly({
  if (input$deaths_scale == "Linear") {
    
    p <- ggplot(filter(uk_data, Date >= "2020-03-06"), aes(x = Date, y = CumDeaths, group = 1, text = paste0("<b>", comma(CumDeaths, accuracy = 1), "</b> deaths<br>", format(Date, "%d %B")))) +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +  
    geom_line(colour = "#8D2313", size = 1) +
    scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), labels = comma) +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank())
    
    ggplotly(p, tooltip = "text") %>%
      config(displayModeBar = FALSE) %>% 
      layout(xaxis = list(fixedrange = TRUE)) %>%
      layout(yaxis = list(fixedrange = TRUE)) 
    
    } else {
      
      temp <- filter(uk_data, Date >= "2020-03-06") %>% 
        arrange(Date) %>% 
        filter(CumDeaths > 0) %>%
        mutate(days = as.integer(Date - min(Date)),
               doubling_2_day = 1*(1+((2^(1/2))-1))^(days),
               doubling_3_day = 1*(1+((2^(1/3))-1))^(days),
               doubling_1_week = 1*(1+((2^(1/7))-1))^(days))
      
      p <- ggplot(temp, aes(x = Date, y = CumDeaths, group = 1, 
                          text = paste0("<b>", comma(CumDeaths, accuracy = 1), "</b> deaths<br>", format(Date, "%d %B")))) +
        geom_hline(yintercept = 1, size = 0.3, colour = "#333333") +
        geom_line(colour = "#8D2313", size = 1) +
        scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
        scale_y_continuous(trans = "log10", expand = c(0.005, 0.005), labels = comma_format(accuracy = 1)) +
        geom_line(aes(x = Date, y = doubling_2_day), colour = "#bdbdbd", lty = 3) +
        geom_line(aes(x = Date, y = doubling_3_day), colour = "#bdbdbd", lty = 3) +
        geom_line(aes(x = Date, y = doubling_1_week), colour = "#bdbdbd", lty = 3) +
        labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
        theme_minimal() +
        theme(plot.margin = unit(rep(0.5, 4), "cm"),
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank())
      
      ggplotly(p, tooltip = "text") %>%
        config(displayModeBar = FALSE) %>% 
        layout(xaxis = list(fixedrange = TRUE)) %>%
        layout(yaxis = list(fixedrange = TRUE)) %>%
        style(text = "Doubling every 2 days", traces = 3) %>%
        style(text = "Doubling every 3 days", traces = 4) %>% 
        style(text = "Doubling every week", traces = 5)
      
    }

})
```

> **Notes**: Logarithmic scales represent an exponential curve as a straight line. The steeper the line the faster the rate of growth.

Maps {data-navmenu="UK"}
=======================================================================

### Cumulative confirmed cases

```{r}
cumulative_cases <- utla_data %>% 
  mutate(popup = str_c("<strong>", area_name, "</strong><br/><strong>", comma(cumulative_lab_confirmed_cases), "</strong> confirmed cases among ", comma(population), " residents<br/>", format(`Specimen date`, "%d %B")) %>% map(HTML))

renderLeaflet({
  leaflet() %>%
  setView(-2.07811, 53.23497, zoom = 7) %>% 
  addTiles(urlTemplate = "https://{s}.basemaps.cartocdn.com/light_nolabels/{z}/{x}/{y}{r}.png",
           attribution = '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/attributions">CARTO</a> | <a href="https://www.ons.gov.uk/methodology/geography/licences">Contains OS data © Crown copyright and database right (2020)</a> | Data: <a href="https://www.gov.uk/government/publications/covid-19-track-coronavirus-cases" target="_blank">Public Health England</a>') %>%
  addPolygons(data = cumulative_cases, fillColor = "#EAEAEA", fillOpacity = 0.3, weight = 1, color = "#bdbdbd")  %>% 
  addCircleMarkers(data = cumulative_cases, lng = ~long, lat = ~lat, radius = ~cumulative_lab_confirmed_cases/100, fillColor = "#57AACB", fillOpacity = 0.8, weight = 1, color = "#FFFFFF", opacity = 1, label = ~popup, labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto")) %>% 
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  onRender(
    paste0("function(el, x) {$('head').append(","\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'",");}"))
})
```

> **Notes**: 1) The size of the circles are proportional to the number of cumulative confirmed cases in each UTLA 2) Upper Tier Local Authorities (UTLAs) include Counties, Unitary Authorities, Metropolitan Districts and London Boroughs. 3) Confirmed cases for Cornwall and Isles of Scilly are combined.<br/>**Source**: [Public Health England](https://coronavirus.data.gov.uk)

### Age-standardised COVID-19 mortality rates
```{r}
age_standardised_deaths <- age_standardised_deaths %>% 
  mutate(popup = ifelse(!is.na(rate), str_c("<strong>", area_name, "</strong><br/><strong>", rate, "</strong> deaths per 100,000 people") %>% map(HTML), str_c("<strong>", area_name, "</strong><br/>", "No data available") %>% map(HTML)))

renderLeaflet({
  breaks <- classIntervals(age_standardised_deaths$rate, n = 5, style = "jenks")$brks
  pal <- colorBin(palette = "Reds", domain = NULL, bins = breaks, na.color = "#D7D8D7")
  
  leaflet(data = filter(age_standardised_deaths, !is.na(rate))) %>%
    setView(-2.07811, 53.23497, zoom = 7) %>% 
    addTiles(urlTemplate = "https://{s}.basemaps.cartocdn.com/light_nolabels/{z}/{x}/{y}{r}.png", attribution = '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/attributions">CARTO</a> | <a href="https://www.ons.gov.uk/methodology/geography/licences">Contains OS data © Crown copyright and database right (2020)</a> | Data: <a href="https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/deaths/datasets/deathsinvolvingcovid19bylocalareaanddeprivation" target="_blank">Office for National Statistics</a>') %>%
    addPolygons(data = filter(age_standardised_deaths, is.na(rate)), fillColor = "#D7D8D7", fillOpacity = 1, weight = 0.5, color = "#bdbdbd", label = ~popup, labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto"), highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE)) %>% 
    addPolygons(fillColor = ~pal(rate), fillOpacity = 0.8, smoothFactor = 0.5, stroke = TRUE, weight = 0.5, color = "#bdbdbd", opacity = 1, label = ~popup, labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto"), highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE)) %>% 
    addLegend(pal = pal, values = ~rate, opacity = 0.7, title = "Rate per 100,000 people<br/><small>1 March - 17 April 2020</small>", position = "bottomright") %>% 
    addFullscreenControl() %>% 
    addResetMapButton() %>% 
    onRender(
      paste0("function(el, x) {$('head').append(","\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'",");}"))
})
```

> **Notes**: Age-standardised mortality rates adjust for the size and age structure of the resident population.<br/>**Source**: [Office for National Statistics ](https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/deaths/datasets/deathsinvolvingcovid19bylocalareaanddeprivation)

International 
=======================================================================

Row
-------------------------------------

### Cumulative number of deaths
```{r}
renderPlotly({
  p <- ggplot(country_data, aes(x = days, y = total_deaths, color = countriesAndTerritories, label = countriesAndTerritories, group = countriesAndTerritories, text = paste0("<b>", countriesAndTerritories, "</b><br><b>", days, "</b> days", "</br><b>", comma(total_deaths), "</b> deaths"))) + 
  geom_hline(yintercept = 100, size = 0.3, colour = "#333333") +
  geom_line(size = 0.6) + 
  scale_colour_manual(values = c("France" = "#52BCA3",  "Italy" = "#99C945", "Germany" = "#CC61B0","Netherlands" = "#24796C", "Spain" = "#2F8AC4", "Sweden" = "#764E9F", "UK" = "#ED645A")) +
  scale_x_continuous(expand = expansion(add = c(0,8))) +
  scale_y_continuous(trans = "log10", breaks = c(100, 200, 500, 1000, 2000, 5000, 10000, 20000, 50000), limits = c(100,NA), expand = expansion(add = c(0,0.1)), labels = comma_format(accuracy = 1)) +
  labs(x = "Days since 100th death", y = NULL, title = NULL, subtitle = NULL, caption = NULL) + 
  theme_minimal() +
  theme(plot.margin = unit(rep(0.5, 4), "cm"),
        panel.grid.minor = element_blank()) +
  guides(colour = FALSE) +
  coord_cartesian(clip = "off")
  
  ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>%
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE)) 

})
```

> **Notes**: 1) Countries have different methods of recording COVID-19 deaths 2) The chart is inspired by the [Financial Times' coronavirus trajectory tracker](https://www.ft.com/coronavirus-latest) which is built and updated by [\@jburnmurdoch](https://twitter.com/jburnmurdoch).<br/>**Source**: [European Centre for Disease Prevention and Control ](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide)

### COVID-19 Government Response Stringency Index
```{r}
renderPlot({
  ggplot(stringency_index,  aes(x = Date, y = Stringency, colour = Country, label = Country, group = Country)) +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_line(size = 1) +
    geom_text_repel(data = ungroup(filter(group_by(stringency_index, Country), Date == max(Date))), aes(x = Date, y = Stringency), fontface = "bold", segment.color = NA, nudge_y = 1.4) +
    scale_colour_manual(values = c("France" = "#52BCA3",  "Italy" = "#99C945", "Germany" = "#CC61B0","Netherlands" = "#24796C", "Spain" = "#2F8AC4", "Sweden" = "#764E9F", "UK" = "#ED645A")) +
    scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), limits = c(0,100)) +
    labs(x = NULL, y = "Stringency Index (0 = no measures)", title = NULL, subtitle = NULL, caption = NULL) + 
    theme_minimal(base_size = 16) +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          strip.text.x = element_blank(),
          axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(colour = FALSE) +
    facet_wrap(~Country, nrow = 1)

}, execOnResize = TRUE)    
```

> **Notes**: The [Oxford COVID-19 Government Response Tracker (OxCGRT)](www.bsg.ox.ac.uk/covidtracker) compares government responses to the coronavirus pandemic by creating a composite index of different categories of response. A score of 0 on the Index indicates no measures have been taken.<br/>**Source**: [Blavatnik School of Government, Oxford University](https://www.bsg.ox.ac.uk/research/research-projects/oxford-covid-19-government-response-tracker).

About {data-icon="fa-info-circle"}
=======================================================================

### 

```{r}
includeHTML("about.html")
```

