---
title: "COVID-19 monitor"
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: columns
    logo: logo.png
    source_code: https://github.com/traffordDataLab/covid-19
    social: ["twitter"]
    mathjax: NULL
runtime: shiny
---

```{css, echo = FALSE}
@import url('https://fonts.googleapis.com/css?family=Open+Sans');

@font-face {
  font-family: 'Open Sans', sans-serif;
}

div {
  font-family: 'Open Sans', sans-serif;
}

h1, h2, h3, h4, h5, h6, .header {
  font-family: 'Open Sans', sans-serif;
}

a {
  color: #046dc3;
  text-decoration: none;
}

a:hover {
  color: #046dc3;
  text-decoration: underline;
}

div.about { 
  font-size: 1.1em;
  font-family: 'Open Sans', sans-serif;
  padding: 20px;
}

.navbar-logo {
  max-width:48px;
  max-height:48px;
}

.tbl {
  font-size: 14px;
  line-height: 18px;
}

/* Hide selection checkboxes in table */
.rt-select-input{
  opacity: 0;
}

/* Change map pointer to grab */
.leaflet-interactive {
    cursor: grab;
}
```

```{r setup, include=FALSE}
library(flexdashboard) ; library(shiny) ; library(tidyverse) ; library(sf) ; library(htmlwidgets) ; library(htmltools) ; library(leaflet) ; library(leaflet.extras) ; library(reactable) ; library(lubridate) ; library(zoo) ; library(scales)

source("global.R")

df <- cases %>% 
  mutate(period = case_when(
    date >= max(date)-days(11) & date <= max(date)-days(5) ~ "current_week",
    date >= max(date)-days(18) & date <= max(date)-days(12) ~ "previous_week"
  )) %>% 
  filter(!is.na(period)) %>% 
  select(-date) %>%
  group_by(area_code, area_name, period) %>% 
  summarise(total_cases = sum(new_cases)) %>% 
  pivot_wider(names_from = period, values_from = total_cases) %>% 
  select(area_code, area_name, previous_week, current_week) %>%
  left_join(population, by = "area_code") %>% 
  mutate(previous_week_rate = round(previous_week/population*100000,1),
         current_week_rate = round(current_week/population*100000,1),
         change = current_week-previous_week,
         change_rate = round(change/population*100000,1)) %>% 
  ungroup() 

# table data
tbl <- select(df, area_name, previous_week_rate, current_week_rate, change_rate)

# map data
sf <- ltla %>% 
  left_join(df, by = "area_code") %>% 
  mutate(popup = str_c("<strong>", area_name, "</strong><br/>", paste(format(max(cases$date)-days(11),'%d %B'),"-", format(max(cases$date)-days(5),'%d %B')),":<strong> ", current_week, "</strong><br/>", paste(format(max(cases$date)-days(18),'%d %B'),"-", format(max(cases$date)-days(12),'%d %B')), ":<strong> ", previous_week, "</strong><br/>Change:<strong> ", change, "</strong>") %>% map(HTML),
         colour = case_when(change > 0 ~ "#C71E1D", change < 0 ~ "#00846C", TRUE ~ "#777"))
```

```{r, reactive}
selected <- reactive(getReactableState("table", "selected"))
```

Dashboard
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

###
```{r}
fillCol(flex = c(NA, 1),
  HTML("<p style = 'color: #707070;font-size: 1.1em;font-weight: bold;margin-left: 15px;'>Confirmed cases per 100,000</p>"),
  reactableOutput("table", height = "100%")
)

output$table <- renderReactable({
  
  reactable(tbl, 
            selection = "single", 
            onClick = "select", 
            highlight = TRUE, 
            rowStyle = list(cursor = "pointer"),
            paginationType = "simple",
            height = "auto",
            compact = TRUE,
            borderless = FALSE,
            wrap = FALSE,
            resizable = TRUE,
            searchable = TRUE,
            defaultSorted = "current_week_rate",
            defaultSortOrder = "desc",
            defaultColDef = colDef(align = "left"),
            theme = reactableTheme(
              rowSelectedStyle = list(backgroundColor = "#eee"),
              searchInputStyle = list(width = "100%")),
            language = reactableLang(
              searchPlaceholder = "Search local authorities...",
              noData = "No local authorities found"),
            columns = list(
              area_name = colDef(name = "Local authority",
                                 align = "left"),
              previous_week_rate = colDef(header =  paste(format(max(cases$date)-days(18),'%d %B'),"-", format(max(cases$date)-days(12),'%d %B')),
                                          align = "left"),
              current_week_rate = colDef(header = paste(format(max(cases$date)-days(11),'%d %B'),"-", format(max(cases$date)-days(5),'%d %B')), 
                                         align = "left"),
              change_rate = colDef(name = "Change",
                                   style = function(value) {
                                     if (value > 0) {
                                       color <- "#C71E1D"
                                       } else if (value < 0) {
                                         color <- "#00846C"
                                         } else {
                                           color <- "#777"
                                           }
                                     list(color = color, fontWeight = "bold")
                                     },
                                     align = "left"))
            )
  
  })
```

###
```{r}
uiOutput("chartUI")

output$chartUI <- renderUI({
  id <- tbl[selected(), ]$area_name
  
  if (is.null(selected())) {
    div(HTML("Please <strong>select a local authority</strong> from the table to show a chart of daily confirmed cases")) 
    } else {
      fillCol(flex = c(NA, 1),
              HTML(paste0("<p style = 'color: #707070;font-size: 1.1em;font-weight: bold;margin-left: 15px;'>Daily confirmed cases in " , id, "</p>")),
              plotOutput("chart")
      )
      }
  })

output$chart <- renderPlot({
  
  id <- tbl[selected(), ]$area_name
  selected_cases <- filter(cases, area_name == id) %>% 
      mutate(ma_cases = rollmean(new_cases, 7, align = "left", fill = NA))
  
  ggplot() +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_col(data = selected_cases, aes(x = date, y = new_cases), fill = "#bdbdbd", colour = NA, width = 0.8, alpha = 0.7) +
    geom_line(data = selected_cases, aes(x = date, y = ma_cases, colour = "ma_cases"), size = 1, alpha = 1) +
    scale_colour_manual(values = c("ma_cases" = "#212121"), name = NULL, labels = "7-day average") +
    scale_x_date(date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), breaks = function(x) unique(
        floor(pretty(seq(0, (max(x) + 1) * 1.1)))), position = "right") +
    labs(x = NULL, y = NULL, title = NULL, NULL,
       caption = "Source: Public Health England") +
    theme_minimal(base_size = 14) +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          plot.caption = element_text(colour = "grey60", margin = margin(t = 20, b = -10)),
          legend.position = "top",
          legend.justification = "left",
          legend.text = element_text(size = 12))
  
}, execOnResize = TRUE) 
```

Column {data-width=500}
-----------------------------------------------------------------------

###
```{r}
uiOutput("mapUI")

output$mapUI <- renderUI({
  if (is.null(selected())) {
    div(HTML("Please <strong>select a local authority</strong> from the table to visualise it and neighbouring areas on a map")) 
    } else {
      fillCol(flex = c(NA, 1),
              HTML(paste0("<p style = 'color: #707070;font-size: 1.1em;font-weight: bold;margin-left: 15px;'>Confirmed cases during week ending ", format(max(pull(filter(cases, date == max(date)-days(5)),date)), '%d %B'),"</p><br/>")),
              leafletOutput("map")
      )
      }
  })

zoom_to <- reactive({
  id <- tbl[selected(), ]$area_name
  filter(sf, area_name == id) 
  })

output$map <- renderLeaflet({
  
    leaflet(data = sf, options = leafletOptions(zoomControl = FALSE)) %>%
    setView(lng = zoom_to()$long, lat = zoom_to()$lat, zoom = 10) %>% 
    addTiles(urlTemplate = "https://{s}.basemaps.cartocdn.com/light_only_labels/{z}/{x}/{y}{r}.png", attribution = '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/attributions">CARTO</a> | <a href="https://www.ons.gov.uk/methodology/geography/licences">Contains OS data © Crown copyright and database right (2020)</a>') %>%
    addPolygons(data = ltla, fillColor = "transparent", weight = 1, color = "#bdbdbd", opacity = 1) %>% 
    addPolygons(data = zoom_to(), fillColor = "#bdbdbd", weight = 1, color = "#bdbdbd", opacity = 1) %>% 
    addCircleMarkers(lng = ~long, lat = ~lat, radius = ~sqrt(current_week), fillColor = ~colour, fillOpacity = 0.3, weight = 1, color = ~colour, opacity = 1, label = ~popup, labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto")) %>% 
    onRender("function(el, t) {var myMap = this;myMap._container.style['background'] = '#ffffff';}", paste0("function(el, x) {$('head').append(","\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'",");}"))
  
  })
```

About {data-icon="fa-info-circle"}
=======================================================================

### 

```{r}
includeHTML("about.html")
```

<script>
  // Changes to improve accessibility
  var el;
  
  // Add document language
  el = document.getElementsByTagName('html')[0];
  el.setAttribute('lang', 'en-GB');
  
  // Alt text for Trafford Council logo and Shiny loading icon
  el = document.getElementsByTagName('img');
  for (var i = 0; i < el.length; i++) {
    if (el[i].className == 'navbar-logo') {
      el[i].setAttribute('alt', 'Trafford Council logo');
    }
    else {
      el[i].setAttribute('alt', '');
    }
  }
  
  // Remove the empty navbar button
  el = document.getElementById('navbar-button');
  el.parentNode.removeChild(el);
  
  // Fix empty Twitter link (empty because it only has an icon and no other content)
  var cbTwitter = setInterval(function(){
    try {
      el = document.getElementsByClassName('fa fa-twitter')[0];
      el.parentNode.setAttribute('aria-label', 'Share this app on Twitter');
      clearInterval(cbTwitter);
    }
    catch(e) {}   // do nothing, wait for next call
  }, 100);
</script>