---
title: "COVID-19 UK monitor"
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: rows
    vertical_layout: fill
    logo: logo.png
    source_code: https://github.com/traffordDataLab/covid-19
    social: ["twitter"]
    css: styles.css
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard) ; library(shiny) ; library(tidyverse) ; library(httr) ; library(htmlwidgets) ; library(htmltools) ; library(scales) ; library(plotly) ; library(ggrepel)

source("load_data.R")
```

Cases {data-navmenu="UK"}
=======================================================================

Row
-------------------------------------

### <strong>New confirmed UK cases</strong><br/><small>as of `r format(max(uk_data$Date), '%d %B %Y')`</small>
```{r}
valueBox(comma(pull(filter(uk_data, Date == max(Date)), NewCases)), color = "#39809E")
```

### <strong>Total confirmed UK cases</strong><br/><small>as of `r format(max(uk_data$Date), '%d %B %Y')`</small>
```{r}
valueBox(comma(max(uk_data$CumCases)), color = "#39809E")
```

Row
-------------------------------------

### Daily confirmed cases
```{r}
renderPlotly({
  p <- ggplot(uk_data, aes(x = Date, y = NewCases, text = paste0("<b>", comma(NewCases, accuracy = 1), "</b> cases<br>", format(Date, "%d %B")))) +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_col(fill = "#39809E") +
    scale_x_date(expand = c(0.005, 0.005), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), labels = comma) +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank()) 
  
  ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>% 
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE))  
})
```

> <br/>**Notes**: 1) Confirmed cases are recorded up to 9am on `r format(max(uk_data$Date), '%d %B %Y')`. 2) There is likely to be a lag of a few days between initial testing and the reporting of test results. 3) The number of *confirmed* cases is likely to be much lower than the *total* number of cases because not all people who report symptoms of coronavirus are being tested.<br/>**Source**: [European Centre for Disease Prevention and Control ](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide)
 
### Cumulative confirmed cases
```{r}
fillCol(flex = c(NA, 1),
        radioButtons("cases_scale", label = NULL, choices = c("Linear", "Logarithmic"), selected = "Linear", inline = TRUE),
        plotlyOutput("cases_plot", height = "100%")
        )

output$cases_plot <- renderPlotly({
  p <- ggplot(uk_data, aes(x = Date, y = CumCases, group = 1, text = paste0("<b>", comma(CumCases, accuracy = 1), "</b> cases<br>", format(Date, "%d %B")))) +
    geom_line(colour = "#39809E", size = 1) +
    scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank())
  
  if (input$cases_scale == "Linear") {
    
    p <- p + geom_hline(yintercept = 0, size = 0.3, colour = "#333333") + scale_y_continuous(expand = c(0.005, 0.005), labels = comma)
    
    } else {
      
      p <- p + geom_hline(yintercept = 1, size = 0.3, colour = "#333333") + scale_y_continuous(trans = "log10", expand = c(0.005, 0.005), labels = comma_format(accuracy = 1))
      
    }
  
  ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>% 
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE)) 
  
})
```

> <br/>**Notes**: Logarithmic scales represent an exponential curve as a straight line. The steeper the line the faster the rate of growth.<br/>**Source**: [European Centre for Disease Prevention and Control ](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide)

Deaths {data-navmenu="UK"}
=======================================================================

Row
-------------------------------------

### <strong>New confirmed UK deaths</strong><br/><small>as of `r format(max(uk_data$Date), '%d %B %Y')`</small>
```{r}
valueBox(comma(pull(filter(uk_data, Date == max(Date)), NewDeaths)), color = "#8D2313")
```

### <strong>Total confirmed UK deaths</strong><br/><small>as of `r format(max(uk_data$Date), '%d %B %Y')`</small>
```{r}
valueBox(comma(max(uk_data$CumDeaths)), color = "#8D2313")
```

Row
-------------------------------------

### Daily confirmed deaths
```{r}
renderPlotly({
  p <- ggplot(filter(uk_data, Date >= "2020-03-06"), aes(x = Date, y = NewDeaths, text = paste0("<b>", comma(NewDeaths, accuracy = 1), "</b> deaths<br>", format(Date, "%d %B")))) +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_col(fill = "#8D2313") +
    scale_x_date(expand = c(0.005, 0.005), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), labels = comma) +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank()) 
  
  ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>% 
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE)) 
})
```

> <br/>**Notes**: 1) Deaths are recorded up to 5pm on `r format(max(uk_data$Date)-1, '%d %B %Y')`. 2) The date of a death refers to the day the death was reported not the actual date of the death. 3) The number of deaths only includes those who tested positive for coronavirus.<br/>**Source**: [European Centre for Disease Prevention and Control ](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide)

### Cumulative confirmed deaths
```{r}
fillCol(flex = c(NA, 1),
        radioButtons("deaths_scale", label = NULL, choices = c("Linear", "Logarithmic"), selected = "Linear", inline = TRUE),
        plotlyOutput("deaths_plot", height = "100%")
        )

output$deaths_plot <- renderPlotly({
  if (input$deaths_scale == "Linear") {
    
    p <- ggplot(filter(uk_data, Date >= "2020-03-06"), aes(x = Date, y = CumDeaths, group = 1, text = paste0("<b>", comma(CumDeaths, accuracy = 1), "</b> deaths<br>", format(Date, "%d %B")))) +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +  
    geom_line(colour = "#8D2313", size = 1) +
    scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), labels = comma) +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank())
    
    ggplotly(p, tooltip = "text") %>%
      config(displayModeBar = FALSE) %>% 
      layout(xaxis = list(fixedrange = TRUE)) %>%
      layout(yaxis = list(fixedrange = TRUE)) 
    
    } else {
      
      temp <- filter(uk_data, Date >= "2020-03-06") %>% 
        arrange(Date) %>% 
        filter(CumDeaths > 0) %>%
        mutate(days = as.integer(Date - min(Date)),
               doubling_2_day = 1*(1+((2^(1/2))-1))^(days),
               doubling_3_day = 1*(1+((2^(1/3))-1))^(days),
               doubling_1_week = 1*(1+((2^(1/7))-1))^(days))
      
      p <- ggplot(temp, aes(x = Date, y = CumDeaths, group = 1, 
                          text = paste0("<b>", comma(CumDeaths, accuracy = 1), "</b> deaths<br>", format(Date, "%d %B")))) +
        geom_hline(yintercept = 1, size = 0.3, colour = "#333333") +
        geom_line(colour = "#8D2313", size = 1) +
        scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
        scale_y_continuous(trans = "log10", expand = c(0.005, 0.005), labels = comma_format(accuracy = 1)) +
        geom_line(aes(x = Date, y = doubling_2_day), colour = "#bdbdbd", lty = 3) +
        geom_line(aes(x = Date, y = doubling_3_day), colour = "#bdbdbd", lty = 3) +
        geom_line(aes(x = Date, y = doubling_1_week), colour = "#bdbdbd", lty = 3) +
        labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL) +
        theme_minimal() +
        theme(plot.margin = unit(rep(0.5, 4), "cm"),
              panel.grid.major.x = element_blank(),
              panel.grid.minor = element_blank())
      
      ggplotly(p, tooltip = "text") %>%
        config(displayModeBar = FALSE) %>% 
        layout(xaxis = list(fixedrange = TRUE)) %>%
        layout(yaxis = list(fixedrange = TRUE)) %>%
        style(text = "Doubling every 2 days", traces = 3) %>%
        style(text = "Doubling every 3 days", traces = 4) %>% 
        style(text = "Doubling every week", traces = 5)
      
    }

})
```

> <br/>**Notes**: Logarithmic scales represent an exponential curve as a straight line. The steeper the line the faster the rate of growth.<br/>**Source**: [European Centre for Disease Prevention and Control ](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide)

International 
=======================================================================

Row
-------------------------------------

### Cumulative number of deaths
```{r}
renderPlotly({
  p <- ggplot(country_data, aes(x = days, y = total_deaths, color = countriesAndTerritories, label = countriesAndTerritories, group = countriesAndTerritories, text = paste0("<b>", countriesAndTerritories, "</b><br><b>", days, "</b> days", "</br><b>", comma(total_deaths), "</b> deaths"))) + 
    geom_hline(yintercept = 100, size = 0.3, colour = "#333333") +
    geom_line(size = 0.6) + 
    geom_text(data = ungroup(filter(group_by(country_data,countriesAndTerritories),days == max(days))), aes(x = days, y = total_deaths), fontface = "bold", check_overlap = TRUE, nudge_x = 0.3, nudge_y = 0.03) +
    scale_colour_manual(values = c("China" = "#E58606", "France" = "#52BCA3",  "Italy" = "#99C945", "Germany" = "#CC61B0","Netherlands" = "#24796C", "South_Korea" = "#DAA51B", "Spain" = "#2F8AC4", "Sweden" = "#764E9F", "UK" = "#ED645A", "USA" = "#CC3A8E")) +
    scale_x_continuous(expand = expansion(add = c(0,1))) +
    scale_y_continuous(trans = "log10", breaks = c(100, 200, 500, 1000, 2000, 5000, 10000, 20000, 50000), limits = c(100,NA), expand = c(0.005, 0.005), labels = comma_format(accuracy = 1)) +
    labs(x = "Days since 100th death", y = NULL, title = NULL, subtitle = NULL, caption = NULL) + 
    theme_minimal() +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.minor = element_blank()) +
    guides(colour = FALSE)
  
  p <- ggplotly(p, tooltip = "text") %>%
    config(displayModeBar = FALSE) %>%
    layout(xaxis = list(fixedrange = TRUE)) %>%
    layout(yaxis = list(fixedrange = TRUE)) %>% 
    hide_legend()
  
  p$x$data[[12]]$hoverinfo <- "none"
  p$x$data[[13]]$hoverinfo <- "none"
  p$x$data[[14]]$hoverinfo <- "none"
  p$x$data[[15]]$hoverinfo <- "none"
  p$x$data[[16]]$hoverinfo <- "none"
  p$x$data[[17]]$hoverinfo <- "none"
  p$x$data[[18]]$hoverinfo <- "none"
  p$x$data[[19]]$hoverinfo <- "none"
  p$x$data[[20]]$hoverinfo <- "none"
  p$x$data[[21]]$hoverinfo <- "none"
  
  p
})
```

> <br/>**Notes**: The number of days for China have been truncated to better show the change in the rate of growth amongst other countries. The chart is inspired by the [Financial Times' coronavirus trajectory tracker](https://www.ft.com/coronavirus-latest) which is built and updated by [\@jburnmurdoch](https://twitter.com/jburnmurdoch).<br/>**Source**: [European Centre for Disease Prevention and Control ](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide)

### COVID-19 Government Response Stringency Index
```{r}
renderPlot({
  ggplot(stringency_index,  aes(x = Date, y = Stringency, colour = Country, label = Country, group = Country)) +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_line(size = 1) +
    geom_text_repel(data = ungroup(filter(group_by(stringency_index, Country), Date == max(Date))), aes(x = Date, y = Stringency), fontface = "bold", segment.color = NA, nudge_y = 1.4) +
    scale_colour_manual(values = c("Germany" = "#CC61B0", "Italy" = "#99C945", "Spain" = "#2F8AC4", "Sweden" = "#764E9F", "UK" = "#ED645A")) +
    scale_x_date(expand = c(0.005, 0.4), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), limits = c(0,100)) +
    labs(x = NULL, y = "Stringency Index (0 = no measures)", title = NULL, subtitle = NULL, caption = NULL) + 
    theme_minimal(base_size = 16) +
    theme(plot.margin = unit(rep(0.5, 4), "cm"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          strip.text.x = element_blank(),
          axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(colour = FALSE)

}, execOnResize = TRUE)    
```

> <br/>**Notes**: The [Oxford COVID-19 Government Response Tracker (OxCGRT)](www.bsg.ox.ac.uk/covidtracker) compares government responses to the coronavirus pandemic by creating a composite index of different categories of response. A score of 0 on the Index indicates no measures have been taken.<br/>**Source**: [Blavatnik School of Government, Oxford University](https://www.bsg.ox.ac.uk/research/research-projects/oxford-covid-19-government-response-tracker).

About {data-icon="fa-info-circle"}
=======================================================================

### 

```{r}
includeHTML("about.html")
```

