---
title: "COVID-19 monitor"
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    orientation: columns
    logo: logo.png
    source_code: https://github.com/traffordDataLab/covid-19
    social: ["twitter"]
    mathjax: NULL
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard) ; library(shiny) ; library(tidyverse) ; library(sf) ; library(htmlwidgets) ; library(htmltools) ; library(leaflet) ; library(leaflet.extras) ; library(reactable) ; library(lubridate) ; library(zoo) ; library(scales)

source("global.R")

df <- cases %>% 
  mutate(period = case_when(
    date >= max(date)-days(8) & date <= max(date)-days(2) ~ "current_week",
    date >= max(date)-days(15) & date <= max(date)-days(9) ~ "previous_week"
  )) %>% 
  filter(!is.na(period)) %>% 
  select(-date) %>%
  group_by(area_code, area_name, period) %>% 
  summarise(total_cases = sum(new_cases)) %>% 
  pivot_wider(names_from = period, values_from = total_cases) %>% 
  select(area_code, area_name, previous_week, current_week) %>%
  left_join(population, by = "area_code") %>% 
  mutate(previous_week_rate = round(previous_week/population*100000,1),
         current_week_rate = round(current_week/population*100000,1),
         change = current_week-previous_week,
         change_rate = round(change/population*100000,1)) %>% 
  ungroup() 

# table data
tbl <- df %>% 
  arrange(desc(current_week_rate)) %>% 
  mutate(rank = rank(desc(current_week_rate), ties.method = "min")) %>% 
  select(rank, area_name, previous_week_rate, current_week_rate, change_rate)
  
# map data
sf <- msoa %>% 
  left_join(msoa_cases, by = "msoa11cd") %>% 
  mutate(popup = ifelse(n == 0,
                        str_c("<strong>", msoa11hclnm, "</strong><br/><em>", lad19nm, "</em><br/>0-2 cases") %>% map(HTML),
                        str_c("<strong>", msoa11hclnm, "</strong><br/><em>", lad19nm, "</em><br/><strong>", n, "</strong> cases") %>% map(HTML)))
```

```{r, reactive}
selected <- reactive(getReactableState("table", "selected"))

observe({
  if (!is.null(selected())) {
    id <- tbl[selected(), ]$area_name
  
    # construct alt text string for the ggplot img tag and send to Javascript handler to add
    altText <- paste0("Plot of daily cases and 7 day average in ", id, ". Source: Public Health England.")
    session$sendCustomMessage("altTextHandler", altText)
  }
})
```

Dashboard
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

###
```{r}
fillCol(flex = c(NA, 1),
  HTML("<p style = 'color: #707070;font-size: 1.1em;font-weight: bold;margin-left: 15px;'>Weekly confirmed cases per 100,000</p>"),
  reactableOutput("table", height = "100%")
)

output$table <- renderReactable({
  
  reactable(tbl, 
            selection = "single", 
            onClick = "select", 
            highlight = TRUE, 
            paginationType = "simple",
            height = "auto",
            compact = TRUE,
            borderless = FALSE,
            wrap = FALSE,
            resizable = TRUE,
            searchable = TRUE,
            defaultSorted = "rank",
            defaultSortOrder = "asc",
            defaultColDef = colDef(align = "left"),
            theme = reactableTheme(
              rowSelectedStyle = list(color = "rgba(255, 255, 255, 1)", backgroundColor = "rgba(57, 128, 158, 0.7)"),
              searchInputStyle = list(width = "100%")),
            language = reactableLang(
              searchPlaceholder = "Search local authorities...",
              noData = "No local authorities found"),
            rowStyle = list(cursor = "pointer"),
            columns = list(
              rank = colDef(name = "Rank", align = "left"),
              area_name = colDef(name = "Local authority", align = "left"),
              previous_week_rate = colDef(header =  paste(format(max(cases$date)-days(15),'%d %b'),"-", format(max(cases$date)-days(9),'%d %b')), align = "left"),
              current_week_rate = colDef(header = paste(format(max(cases$date)-days(8),'%d %b'),"-", format(max(cases$date)-days(2),'%d %b')), align = "left"),
              change_rate = colDef(name = "Change", align = "left"))
            )
  
  })
```

###
```{r}
uiOutput("chartUI")

output$chartUI <- renderUI({
  id <- tbl[selected(), ]$area_name
  
  if (is.null(selected())) {
    div(HTML("Please <strong>select a local authority</strong> from the table to show a chart of daily confirmed cases")) 
    } else {
      fillCol(flex = c(NA, 1),
              HTML(paste0("<p style = 'color: #707070;font-size: 1.1em;font-weight: bold;margin-left: 15px;'>Daily confirmed cases in " , id, "</p>")),
              plotOutput("chart")
      )
      }
  })

output$chart <- renderPlot({
  
  id <- tbl[selected(), ]$area_name
  selected_cases <- filter(cases, area_name == id, 
                           # exclude incomplete cases
                           date <= max(date)-days(2)) %>% 
      mutate(ma_cases = rollmean(new_cases, 7, align = "center", fill = NA))
  
  ggplot() +
    geom_hline(yintercept = 0, size = 0.3, colour = "#333333") +
    geom_col(data = selected_cases, aes(x = date, y = new_cases), fill = "#39809E", colour = NA, width = 0.8, alpha = 0.7) +
    geom_line(data = selected_cases, aes(x = date, y = ma_cases, colour = "ma_cases"), size = 1, alpha = 1) +
    scale_colour_manual(values = c("ma_cases" = "#757575"), name = NULL, labels = "7-day average") +
    scale_x_date(breaks = c(min(selected_cases$date), max(selected_cases$date)), date_labels = "%d-%b") +
    scale_y_continuous(expand = c(0.005, 0.005), breaks = function(x) unique(
        floor(pretty(seq(0, (max(x) + 1) * 1.1)))), position = "right") +
    labs(x = NULL, y = NULL, title = NULL, NULL,
       caption = "Source: Public Health England") +
    theme_minimal(base_size = 14) +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          plot.caption = element_text(colour = "grey60", margin = margin(t = 20, b = -6)),
          legend.position = "top",
          legend.justification = "left",
          legend.text = element_text(size = 12),
          axis.ticks.x = element_line(colour = 'black', size = 0.5))
  
}, execOnResize = TRUE) 
```

Column {data-width=500}
-----------------------------------------------------------------------

###
```{r}
uiOutput("mapUI")

output$mapUI <- renderUI({
  id <- tbl[selected(), ]$area_name
  
  if (is.null(selected())) {
    div(HTML("Please <strong>select a local authority</strong> from the table to show a map of case counts by neighbourhood")) 
    } else {
      fillCol(flex = c(NA, 1),
              HTML(paste0("<p style = 'color: #707070;font-size: 1.1em;font-weight: bold;margin-left: 15px;'>Confirmed cases over the last 7 days in ", id, "</p>")),
              leafletOutput("map")
      )
      }
  })

zoom_to <- reactive({
  id <- tbl[selected(), ]$area_name
  filter(sf, lad19nm == id) 
  })

output$map <- renderLeaflet({
  
    leaflet(data = zoom_to()) %>%
    addTiles(urlTemplate = "https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}{r}.png", attribution = '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/attributions">CARTO</a> | <a href="https://www.ons.gov.uk/methodology/geography/licences">Contains OS data © Crown copyright and database right (2020)</a>') %>%
    addPolygons(fillColor = "transparent", weight = 1, color = "#757575", opacity = 1) %>% 
    addCircleMarkers(lng = ~lon, lat = ~lat, radius = ~sqrt(n), fillColor = "#39809E", fillOpacity = 0.7, weight = 1, color = "#FFFFFF", opacity = 1) %>%
    addPolygons(fillColor = "transparent", stroke = TRUE, weight = 1, color = "transparent", opacity = 1, label = ~popup, labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "bottom", offset = c(0, 25)), highlight = highlightOptions(fillColor = "#39809E", fillOpacity = 0.2, stroke = TRUE, color = "#39809E", weight = 1, bringToFront = TRUE)) %>%
  onRender(paste0("function(el, x) {$('head').append(","\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'",");}"))
  
  })
```

About {data-icon="fa-info-circle"}
=======================================================================

### 

```{r}
includeHTML("about.html")
```

<script>
  // Changes to improve accessibility
  var el;
  
  // Add document language
  el = document.getElementsByTagName('html')[0];
  el.setAttribute('lang', 'en-GB');
  
  // Alt text for Trafford Council logo and Shiny loading icon
  el = document.getElementsByTagName('img');
  for (var i = 0; i < el.length; i++) {
    if (el[i].className == 'navbar-logo') {
      el[i].setAttribute('alt', 'Trafford Council logo');
    }
    else {
      el[i].setAttribute('alt', '');
    }
  }
  
  // Give the empty navbar menu button a descriptive label (this is the button which appears when the screen width is narrow)
  el = document.getElementById('navbar-button');
  el.setAttribute('aria-label', 'toggle menu items list')
  
  // Fix empty Twitter link (empty because it only has an icon and no other content)
  var cbTwitter = setInterval(function(){
    try {
      el = document.getElementsByClassName('fa fa-twitter')[0];
      el.parentNode.setAttribute('aria-label', 'Share this app on Twitter');
      clearInterval(cbTwitter);
    }
    catch(e) {}   // do nothing, wait for next call
  }, 100);
  
  // Make the title next to the logo a <h1> heading
  el = document.getElementsByClassName('navbar-brand')[0];
  var h1 = document.createElement('h1');
  var h1Content = el.replaceChild(h1, el.childNodes[0]);
  h1.appendChild(h1Content);
  
  // Add aria-live attributes to the chart and map elements so that when their contents change it will be announced
  el = document.getElementById('chartUI');
  el.setAttribute('aria-live', 'polite');
  el = document.getElementById('mapUI');
  el.setAttribute('aria-live', 'polite');
  
  /* --- alt text for chart img --- */
  var altTextCache = "";
  
  // Listen for window resizing as this will cause the chart to redraw and lose the alt text
  window.addEventListener('resize', function () {
    var chartContainer = document.getElementById('chartUI');
    
    // We only want to update the alt text if there is a chart present. The following classes are only present if the chart is displayed
    if (chartContainer.childNodes[0].classList == 'flexfill-container flexfill-container-column') {
      var altTextCallback = setInterval(function () {
        // call the fn to update the alt text and clear the callback if successful
        if (updateAltText(altTextCache)) clearInterval(altTextCallback);
      }, 500);
    }
  });
  
  // Receive call from Shiny server with the alt text for the dynamic plot img.
  Shiny.addCustomMessageHandler('altTextHandler', function(altText) {
    altTextCache = altText;   // cache the alt text in case is needs reinstating on window resize
    
    var altTextCallback = setInterval(function () {
        // call the fn to update the alt text and clear the callback if successful
        if (updateAltText(altText)) clearInterval(altTextCallback);
      }, 100);
  });
  
  function updateAltText(text) {
    /*
      Attempt to apply the alt text to the img, which is the first child of <div id='chart'>. 
      If it fails it's usually because Shiny hasn't created either element yet.
      If this happens the interval will call the function repeatedly until we succeed.
    */
    try {
      var plotContainer = document.getElementById('chart');
      plotContainer.firstChild.setAttribute('alt', text);
      return true;
    }
    catch(e) { return false; }
  }
  /* ------------------------------ */
  
  // Add CSS at end of head element within a style tag rather than as a link in the head as this doesn't have the required level of specificity and rather than a css r block as this creates a style tag within the body which is not valid HTML 
  var css = document.createElement('style');
  document.head.appendChild(css);
  css.innerHTML = `
@import url('https://fonts.googleapis.com/css?family=Open+Sans%7CRoboto');

html, body {
  font-family: 'Open Sans', sans-serif;
}

h1, h2, h3, h4, h5, h6 {
  font-family: 'Roboto', sans-serif;
  color: #707070;
}

h1
{
  display: inline-block;
  font-size: 1.25em;
  margin-top: -1em;
  padding-top: 0;
}

a {
  color: #046dc3;
  text-decoration: none;
}

a:hover {
  color: #046dc3;
  text-decoration: underline;
}

div.about { 
  font-size: 1.1em;
  font-family: 'Open Sans', sans-serif;
  padding: 20px;
}

.navbar-logo {
  max-width:48px;
  max-height:48px;
}

#navbar li a
{
  color: #707070;
}

#navbar li a:hover, #navbar li.active a
{
  color: #212121 !important;
}

.tbl {
  font-size: 14px;
  line-height: 18px;
}

/* Hide selection checkboxes in table */
.rt-select-input{
  opacity: 0;
}

/* Change map pointer to grab */
.leaflet-interactive {
    cursor: grab;
}
  `;
</script>